\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{microtype}
\usepackage{parskip}
\usepackage[hang,flushmargin]{footmisc}

\emergencystretch=3em
\hbadness=10000

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=black,
  citecolor=black,
  pdftitle={Tacit: A Confidential, Self-Custodial Asset Protocol on Bitcoin},
  pdfauthor={z0r0z}
}

\title{Tacit: A Confidential, Self-Custodial Asset Protocol on Bitcoin}
\author{z0r0z \\ \small\texttt{https://tacit.finance}}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
\noindent
A purely on-chain version of confidential value, anonymous spend,
shielded recipient addresses, automated market making, and wrapped
Bitcoin would let users hold and exchange programmable assets on
Bitcoin L1 without trusting a federation, sidechain, or bridge. Signatures and rangeproofs solve
part of the problem; the main benefits are lost if a trusted party still
enforces token rules or custodies value. The unlock is a \textbf{composition
of self-custodial primitives}: two Groth16 circuit families (anonymous
unique-spend and amount confidentiality), a Runes-style indexer-validated
meta-protocol, and a fixed-supply tacit asset (TAC). TAC's market value
lets indexer-validated state enforce economic constraints, and its
intrinsic privacy value as a confidential, self-custodial asset
justifies its existence alongside Bitcoin. Any indexer running the specification reaches
the same verdict from chain data alone. Wrapped Bitcoin is locked
cryptographically at a Taproot key derived from a mixer note's own
secret: one leaf, two locks. Amount-granular fungible BTC instead replaces the federation with an
LP-share lien on the canonical AMM pool, with TAC economic
over-collateralization as the bond. Together
these primitives compose into real Bitcoin DeFi as a self-custodial
layer --- confidential assets, native AMM, yield farms, atomic-OTC
marketplace, wrapped BTC --- anchored to Bitcoin L1, recoverable from
each user's private key, and free of federation, custodian, or
operator-set trust. The cryptographic claims hold under standard
assumptions; the economic claims hold conditional on TAC market depth and
an active liquidator population, both enumerated explicitly.
\end{abstract}

\section{Introduction}

Bitcoin solved double-spend over a peer-to-peer network. Doing more than
moving sats --- confidential transfers, anonymous spend, automated market
making, wrapping Bitcoin into programmable form --- has historically required
either off-chain proof exchange (RGB, Taproot Assets), a federated sidechain
(Liquid), a rollup or sidechain with an operator set (Citrea, Botanix), or a
custodial or threshold-bonded wrapper (WBTC, tBTC). Each reintroduces a
trusted party.

What is needed is a layer that (i) keeps every byte of state on Bitcoin L1,
(ii) hides amounts cryptographically, (iii) supports anonymous spend without
a coordinator, (iv) lets Bitcoin be wrapped into a fungible asset without a
federation, and (v) recovers entirely from a private key. Tacit is that
layer --- a composition of self-custodial primitives that delivers DeFi
on Bitcoin L1 without a runtime, custodian, or operator set.

\section{Indexer-validated meta-protocol}

\textbf{Tacit is an indexer-validated layer on Bitcoin L1.} Envelopes
ride inside Bitcoin Taproot script-path witnesses; Bitcoin nodes do
not interpret them. \emph{Indexers} --- open-source clients anyone
can run --- scan confirmed transactions and reach the same verdict
from chain data alone. This is the trust model Runes and
Ordinals already use to underwrite market
value\footnote{Rodarmor, C. \emph{Runes Protocol.} 2024.}: two indexers
seeing the same byte sequence reach the same accept/reject decision,
byte-for-byte. There is no voting, no quorum, no leader. Determinism is
the substrate.

The trust target is therefore not ``indexer determinism in the abstract''
but the \textbf{published specification and its open-source reference
implementations} --- the same shape Bitcoin Core has at L1. Contested
edge cases resolve via SPEC clarification and cross-implementation test
vectors, not by quorum vote. This is a softer property than a federation
and a stronger one than off-chain proof distribution, but it is not a
solved problem in the abstract: it depends on the protocol's open-source
process remaining contestable. \S13 enumerates the limits this implies.

The contribution is to use that substrate for more than tokens. The
same consensus-of-indexers that enforces who-owns-what can also enforce
who-bonded-what and who-equivocated-when, \emph{provided} there exists
a tacit asset the protocol can natively slash. TAC is that asset: a
fair-launched tacit asset with a fixed
$21{,}000{,}000$-base-unit cap\footnote{\texttt{SPEC-WORKER-BOND-AMENDMENT}
--- \emph{Why burn the slashed TAC instead of pooling it}: TAC fixed
cap, burn-on-slash semantics.}, deployed via the same permissionless
\texttt{T\_PETCH}/\texttt{T\_PMINT} mechanism any other auditable-supply
token uses. The protocol uses TAC as the bond layer for cBTC.tac (\S6),
workers (\S9), and bounded governance.

\paragraph{Indexer determinism + a fixed-cap tacit asset together
replace both the federation and the smart-contract VM.} Sidechains
keep the federation (Liquid, ${\sim}11$-of-$15$ multisig); rollups and
Bitcoin L2s keep a VM but inherit an operator set (Citrea, Botanix);
each reintroduces a trust party tacit avoids. Where cryptography can
enforce a property directly --- custody of BTC at L1, amount
confidentiality, anonymous spend --- the protocol uses cryptography.
Where it cannot --- fungibility of wrapped BTC, soft-confirms before
Bitcoin confirmation --- TAC stake fills the gap. Because the cap is
fixed and slashed TAC burns against it, the bonded layer's strength
compounds: each burn permanently shrinks the circulating float against
a fixed cap, so every remaining holder's relative share strengthens
and every future bond stacks against a smaller supply. Slashing schemes
that recycle proceeds into recoverable treasuries, or that coexist with
continuous issuance, do not have this monotone property.

\section{Confidential value}

Every on-chain amount is a Pedersen commitment on
secp256k1\footnote{Pedersen, T. P. \emph{Non-interactive and
information-theoretic secure verifiable secret sharing.} CRYPTO 1991.}:
\[
C \;=\; a \cdot H \;+\; r \cdot G,
\]
where $G$ is Bitcoin's generator, $H$ is a NUMS generator (no known
$\log_{G} H$), $a$ is the amount, and $r$ is a blinding scalar.
Commitments are perfectly hiding and additively homomorphic, so supply
conservation holds across a transaction by checking
\[
\sum_{\text{out}} C \;-\; \sum_{\text{in}} C \;=\; r_{\text{ex}} \cdot G,
\]
where $r_{\text{ex}}$ is the excess. A BIP-340 Schnorr signature\footnote{
Wuille, P.; Nick, J.; Ruffing, T. \emph{BIP-340: Schnorr signatures for
secp256k1.} 2020.} under $r_{\text{ex}} \cdot G$ --- a Mimblewimble-style
\textbf{kernel signature}\footnote{Jedusor, T. E. \emph{Mimblewimble.} 2016;
Poelstra, A. \emph{Mimblewimble.} 2016.} --- proves
$\sum_{\text{out}} a = \sum_{\text{in}} a$ without revealing any $a$.
Range is enforced by one aggregated Bulletproof\footnote{B\"unz, B.; Bootle,
J.; Boneh, D.; Poelstra, A.; Wuille, P.; Maxwell, G. \emph{Bulletproofs:
Short proofs for confidential transactions and more.} IEEE S\&P 2018.}
covering all outputs at $n = 64$ bits, $m \in \{1,2,4,8\}$.

The recipient's blinding and amount-keystream are both derived from
the ECDH shared secret of the sender's private key and the recipient's
public key, with domain-separated HMAC-SHA256. The sender's pubkey appears in the input
witness; the recipient re-derives every credit from their private key and chain
alone. \textbf{No share-link, no sync server.}

\section{Anonymous unique-spend}

Amount privacy alone leaves the address graph visible. Tacit's mixer adds a
Tornado-style shielded pool per $(\mathit{asset\_id}, \mathrm{denom})$ pair.
Deposit appends a Poseidon-hash leaf\footnote{Grassi, L.; Khovratovich, D.;
Rechberger, C.; Roy, A.; Schofnegger, M. \emph{Poseidon: A new hash function
for zero-knowledge proof systems.} USENIX Security 2021.}
$\mathrm{leaf} = \mathrm{Poseidon}_3(s, \nu, \mathrm{denom})$ to a depth-$20$
Merkle tree the indexer maintains. Withdrawal produces a Groth16 proof over
BN254\footnote{Groth, J. \emph{On the size of pairing-based non-interactive
arguments.} EUROCRYPT 2016.} of the statement
\[
\exists\, (s, \nu, \pi) \;:\; \mathrm{leaf}(s,\nu,d) \in \mathrm{root},\;
\mathrm{Poseidon}_1(\nu) = \mathrm{null},\;
r_{\text{leaf}} = \mathrm{Poseidon}_2(s,\nu),
\]
with public inputs
$(\mathrm{root},\,\mathrm{null},\,d,\,r_{\text{leaf}},\,\mathrm{bind})$,
where $\mathrm{bind}$ is SHA-256 over withdrawal-specific data (asset
id, denomination, nullifier, recipient commitment, $r_{\text{leaf}}$)
reduced to a BN254 Fr element. The validator recomputes it externally;
because the circuit constrains $\mathrm{bind}$ as a public input,
Groth16's input-commitment vector binds the proof to that specific
value, and any reuse against a different recomputed tuple fails
verification --- preventing replay or front-running. Indexers reject any reused $\mathrm{null}$. The deposit and withdrawal
are unlinkable inside the anonymity set of unspent leaves. The same
circuit underpins cBTC.zk slot operations (\S5), where leaf-level
privacy holds but each slot's backing BTC UTXO is still a visible
Bitcoin chain-graph element --- protocol-layer privacy without
BTC-layer privacy\footnote{See \texttt{spec/GLOSSARY.md} ---
\emph{Two privacy layers, not one}.}.

\section{cBTC.zk: one leaf, two locks}

BN254 Fr scalars (order ${\approx}2^{254}$) embed into secp256k1's scalar
field (order ${\approx}2^{256}$) without reduction, so $r_{\text{leaf}}\cdot
G_{\text{secp256k1}}$ is a valid Bitcoin public key. We lock real BTC at a
key-path Taproot output
\[
K_{\text{btc}} \;=\; r_{\text{leaf}} \cdot G_{\text{secp256k1}}
\;=\; C \;-\; d\cdot H,
\]
derivable by any indexer from the on-chain Pedersen commitment $C$.
\textbf{The same scalar that proves set membership inside the SNARK is the
spending key of the backing UTXO.} A user who can withdraw the note can
spend the BTC; nobody else can. No federation, no co-signer, no oracle,
no escape script --- one leaf, two locks. We call this asset
\textbf{cBTC.zk}.

\paragraph{Transfer rekeys atomically.} Moving a cBTC.zk slot to a new
holder spends the old $K_{\text{btc}}$ and creates a new
$K_{\text{btc}}' = r_{\text{leaf}}' \cdot G_{\text{secp256k1}}$ in the
same Bitcoin transaction (\texttt{T\_SLOT\_ROTATE},
\texttt{T\_SLOT\_SPLIT}, or \texttt{T\_SLOT\_MERGE}). The new key is
derived from a secret the recipient controls (the sender never learns
it), so the previous holder retains no spending authority over the
backing BTC UTXO. cBTC.zk is
therefore transferable as a first-class slot asset, not merely a
self-held vault note.

\paragraph{Non-upgradability is the price of trustlessness.} The slot
layer has no escape script and no break-glass. The same
cryptography that excludes federations from custody also excludes
recovery if a circuit bug is ever found. A successor pool with a new
circuit ships as a separate $(\mathit{asset\_id},\,\mathrm{denom})$
surface; existing slots stay on their original key and migrate
slot-by-slot through the standard withdraw-and-redeposit flow. This is
the same trade Bitcoin itself makes at L1 --- cryptographic rigidity in
exchange for the absence of an upgrade party.

Any spend of $K_{\text{btc}}$ without a matching $\mathrm{T\_SLOT\_BURN}$
envelope is observable from chain alone (the slot's nullifier is missing
from the consumed set). This $\mathrm{SLASH\_DETECTED}$ event is the
primitive on which later economic constructions rest.

\section{cBTC.tac: bonded fungibility}

cBTC.zk is unit-granular at fixed denominations and trustless. DeFi
expects amount-granular fungible assets. \textbf{cBTC.tac} composes a
cBTC.zk slot with an LP-share lien on the canonical
$(\mathrm{cBTC.zk},\,\mathrm{TAC})$ AMM pool. Minting requires the
depositor to LP both sides; the indexer records a lien on the resulting
LP-share UTXO and mints fungible cBTC.tac equal to the slot's BTC value.
The lien is enforced by the wallet's universal
$\mathrm{commitmentForUtxo}$ resolver: any unauthorized spend is treated
as nonexistent, so AMM swaps, transfers, and farms refuse to consume
liened collateral\footnote{\texttt{SPEC-CBTC-TAC-AMENDMENT} \S5.47 ---
lien model (trustless collateral without covenants).}.

If the depositor rugs the underlying slot
($\mathrm{SLASH\_DETECTED}$), their TAC contribution moves to a shared
insurance pool that compensates all outstanding cBTC.tac holders pro
rata. With initial bond ratio $\geq 2.0$ relative to slot BTC value and
force-close defending at $1.2\times$, the rug is negative expected
value while the liquidator population operates as designed. \textbf{cBTC.tac is
trustless on the BTC anchor side (cryptographic) and over-collateralized
on the fungibility side (economic) --- the same shape as DAI but with no
smart-contract runtime.} This makes cBTC.tac a first-class tacit asset: it
transfers via standard CXFER, swaps in any pool, deposits into the mixer,
and stakes into yield farms. Bitcoin gets a fungible BTC-denominated asset
inside a DeFi stack anchored to L1, without a custodian.

\paragraph{Stress behavior and reflexive collateral.} The construction
is procyclical. TAC's market value derives in part from its role as the
bond, so a stress event that strains cBTC.tac can also strain TAC ---
the collateral is partially \textbf{reflexive}, not merely correlated.
The 2.0$\times$ initial cushion, the permissionless force-close path
triggered when LP-share BTC value falls below 1.2$\times$ the slot, and
the shared insurance pool bound the damage. Non-bond TAC demand (AMM
market-making, payments, governance) is what mitigates the reflexivity
in practice. The trust model under stress is ``rational liquidators arrive
in time,'' the same shape Maker has. Maker collateralizes against ETH's
mature exogenous demand; TAC's parallel demand develops through the
non-bond uses enumerated in \S13 --- worker bonding, governance, AMM
denomination, protocol-fee revenue (\S8), farm rewards, and TAC's role
as a native confidential asset.

\section{Three privacy capabilities, composable and optional}

Privacy is exposed as three orthogonal capabilities, each opt-in at a
different granularity. Users pick the posture that fits their use case
instead of forcing one stance on everyone.

\begin{table}[h]
\centering
\small
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}p{0.20\linewidth} p{0.16\linewidth} p{0.27\linewidth} p{0.27\linewidth}@{}}
\toprule
\textbf{Capability} & \textbf{Default} & \textbf{Technique} & \textbf{What it hides} \\
\midrule
Shielded amount & On (every transfer) &
Pedersen commit $C = aH + rG$, aggregated Bulletproof for $a \in [0,2^{64})$,
kernel signature on the excess; ECDH-derived recipient keystream &
Per-output amounts. Only the public \texttt{burned\_amount} field of
\texttt{T\_BURN} is exposed by design --- for auditability of supply
reductions. \\
Shielded address & Opt-in per receipt &
Blinded-pubkey commit
$\mathrm{commit} = P_{\text{recipient}} + b\cdot G$ with
$b = \mathrm{HMAC}(\mathit{shared}\,\Vert\,\mathit{anchor})$,
$\mathit{shared} = \mathrm{ECDH}(\cdot)$.
Static handle \texttt{tcs1\dots} encodes the recipient pubkey opaquely &
The recipient's on-chain identity. Each receipt lands at a
per-transaction unique \texttt{P2WPKH(hash160(commit))} with no apparent
link to the published handle. Same key-tweak primitive as
BIP-341\footnote{Wuille, P.; Nick, J.; Towns, A. \emph{BIP-341: Taproot.}
2020.} / BIP-352\footnote{\emph{BIP-352: Silent Payments.} 2023.}; no new ceremony. \\
Mixer pool & Opt-in per UTXO &
Tornado-style Poseidon-Merkle commitment with Groth16 proof of
unspent-leaf membership and nullifier reveal (\S4) &
The link between deposit and withdraw. Inside the pool's anonymity set,
which leaf was touched is hidden; both endpoints belong to the same
$(\mathit{asset\_id}, \mathrm{denom})$ surface. \\
\bottomrule
\end{tabular}
\end{table}

\noindent
The three layers are orthogonal because each operates at a different
position in the wire: \textbf{shielded amount} on the commitment value,
\textbf{shielded address} on the output script, \textbf{mixer pool} on
the protocol-tree leaf. A given UTXO can carry any subset
independently, and a wallet can dial privacy up or down per-payment
without giving up recovery.

\paragraph{Practical postures.}
A \emph{merchant} keeps default shielded amounts and a static public
identity for accounting --- balances are private but counterparties
remain identifiable for invoicing and reconciliation. \emph{Payroll}
runs a batched CXFER from one treasury key to $N$ employees, each of
whom publishes a \texttt{tcs1\dots} shielded address: every employee's
on-chain pay line lands at a fresh, unlinkable marker, and the
employee re-derives the tweaked signing key from their wallet seed
alone. A \emph{privacy-conscious user} routes outbound flow through
the mixer pool and receives inbound at a shielded address ---
unlinkable at both endpoints, with full unlinkability inside the
chosen anonymity set. Each posture composes \textbf{the same three
primitives} differently; the protocol itself does not change between
them.

\section{Native AMM and LP-bonded farms}

The AMM follows the same virtual-pool pattern as the mixer: a pool of
$(R_A, R_B, S)$ is \textbf{just numbers} that the indexer reconstructs
from chain. Two trader paths:
\begin{itemize}\setlength\itemsep{2pt}
\item $\mathrm{T\_SWAP\_VAR}$ --- per-trade against the spot curve
$\Delta_{\text{out}} =
R_B\,\gamma\,\Delta / (R_A\,\gamma_d + \gamma\,\Delta)$ ($\gamma$ and
$\gamma_d$ are the per-pool fee numerator and denominator, e.g.\
$\gamma=997$, $\gamma_d=1000$ for a 30~bps pool) using only Pedersen
$+$ Bulletproof $+$ kernel sig. \textbf{No Groth16, no
ceremony, and per-trade $\Delta$ is cleartext on chain} --- the explicit
trade-off for skipping the SNARK. Variable-amount $[Y, X]$ semantics,
immediate settlement. Privacy-sensitive flow routes through
$\mathrm{T\_SWAP\_BATCH}$ instead.
\item $\mathrm{T\_SWAP\_BATCH}$ --- uniform-clearing-price
batch\footnote{Walras, L. \emph{\'El\'ements d'\'economie politique pure.}
1874; Gnosis Protocol uniform-clearing-price batches, 2019; Penumbra
ZSwap, 2023.} over $N$ hidden per-trader amounts, settled by one Groth16
proof. Intra-batch sandwich attacks are structurally impossible because
every trader in a batch clears at the same $P_{\text{clear}}$.
\end{itemize}

LP shares are themselves confidential tacit assets, so they compose with
the mixer (anonymous LP positions) and with cBTC.tac (LP-share-lien
collateral). Yield farms ($\mathrm{T\_LP\_BOND}$) stream reward
emissions to bonded LP shares without a smart contract: the farm
treasury is a virtual indexer-attested quantity; bond receipts are
P2WPKH dust markers keyed to per-bond accrual snapshots; the launcher's
reward asset enters via kernel-sig closure and exits when the validator
credits a fresh reward output at unbond\footnote{
\texttt{SPEC-AMM-FARM-AMENDMENT} --- virtual-treasury MasterChef-style
farms.}. Bootstrap depth is incentivized without an on-chain treasury
or custody role.

\paragraph{Optional protocol-fee skim.} Pools may pin a fee skim at
\texttt{POOL\_INIT}, directing a portion of trade revenue to a
configurable recipient. Canonical pools direct it to the TAC treasury,
so AMM volume accrues structural revenue to TAC holders alongside
organic LP fees. \textbf{TAC is the protocol's DAO-governed
DeFi-native unit} --- canonical to the AMM, bonding, and governance
from launch, not a generic asset adapted to collateral after the
fact. The same trade activity that prices cBTC.tac through TWAP
collateralization also accrues value at the bond layer. With the full
DeFi primitive set on one stack --- fungible wrapped BTC, AMM, LP
yield, farm incentives, snapshot airdrops, future stablecoin variants
--- pool counterparties span TAC, cBTC.tac, and any other tacit asset,
so depth compounds within the protocol's value-accrual surface rather
than fragmenting to external venues. A structural difference from
MakerDAO, where the AMM that prices ETH/DAI runs on external venues
and its fees flow elsewhere.

\section{Validator honesty and bonded diversity}

The channel layer ($\mathrm{T\_INTENT\_ATTEST}$) lets workers offer
${\sim}30\,\text{s}$ soft-confirmations on pending intents ---
pre-broadcast trade commitments that haven't yet settled on Bitcoin.
Limit orders that never match cost zero on-chain; matched orders
amortize into batched settlement. The construct is only useful if
equivocation is costly.

A worker who wants traders to honor its attestations posts a TAC bond via
\texttt{T\_WORKER\_BOND\_OPEN}. The bond UTXO is an ordinary TAC-bearing
UTXO carrying an \textbf{indexer-recorded lien} --- no covenant needed;
tacit-aware wallets refuse to consume liened TAC. Any party who observes
two attestations signed by the same \texttt{worker\_pubkey} at the same
$(\mathit{scope\_id},\,\mathit{height})$ with conflicting
\texttt{intent\_pool\_hash} can submit them as evidence; the indexer
reattributes a configurable bounty fraction to the reporter and
\textbf{burns the remainder}, contracting circulating TAC against its
fixed 21M-base-unit cap. Any observer has positive-EV incentive to
report. The structure mirrors Ethereum's PoS slashing\footnote{Buterin,
V., et al. \emph{Ethereum Proof-of-Stake / Casper FFG.} 2020.} but is
enforced by indexer consensus rather than a smart contract, and
inherits the monotone-deterrent property from \S2.

Workers cannot make invalid envelopes valid; honest competition for
intent flow plus burn-on-equivocation suffices to align worker behavior.
Anyone can run a worker, and the dapp's \textbf{tacit-mesh}
cross-checker turns that diversity into evidence: a worker that quietly
omits an intent from its published list while signing an attestation
committing to the omitted set is caught by a SHA-256 comparison across
$\geq 2$ subscribed workers. The mesh produces evidence; the bond turns
it into a slash.

\paragraph{Bond-class accounting.} Worker bonds and cBTC.tac collateral
share the same TAC float and are both recorded as indexer liens against
the fixed cap. A system-wide governance parameter caps the bonded
fraction of circulating TAC, and burns from either class contract
supply for the other. The classes compete for float but compose
cleanly: one TAC, one cap, one ledger of liens.

\section{Self-custody and recovery}

A wallet recovers full state from its private key and the Bitcoin chain
alone. The derivation paths by primitive:
\begin{itemize}\setlength\itemsep{2pt}
\item \emph{Confidential transfers:} trial-decrypt outputs via ECDH
against the sender pubkey carried in the input witness.
\item \emph{Own etches and mints:} re-derive supply blinding/keystream
from the commit-tx anchor and the wallet's own key.
\item \emph{Permissionless mints and mixer withdrawals:} the opening
$(a, r)$ is published cleartext in the envelope (verified via the
Pedersen equation).
\item \emph{AMM and farm positions:} the public amount
(\texttt{share\_amount}, $\Delta_a$, $\Delta_b$) sits in the envelope;
the chain commitment uses a deterministic blinding derived from the
wallet key plus a per-tx anchor.
\item \emph{Shielded-address receipts:} scan a bounded
\texttt{xferseen} index per asset and re-derive the per-tx tweak $b$
from ECDH and the anchor; the spending key is $sk_{\text{recipient}} +
b \bmod n$.
\end{itemize}

\paragraph{Shielded-pool deposits and cBTC.zk slots.} Each deposit's
$(\mathit{secret}, \nu)$ pair is wallet-derived at deposit time and
consumed at withdraw time. Standard tacit wallets derive the pair
deterministically from the wallet seed (a BIP-32-style sub-key seeds
the per-deposit CSPRNG, anchored on the deposit tx), so recovery
re-derives candidate notes and scans the pool's leaf set for matches
--- privkey-only, same as the rest of the protocol. The protocol
consumes any valid $(\mathit{secret}, \nu)$ pair without distinguishing
how it was generated, so a wallet may also choose per-deposit CSPRNG
with out-of-band note backup (the Tornado / Privacy Pools posture);
both paths produce identical on-chain leaves and identical withdraws.
\textbf{Fungible cBTC.tac balances} held as ordinary tacit-asset UTXOs
recover from the wallet key alone like any other tacit asset.

A user who loses every device and every backup except their private
key reconstructs every asset class from the chain alone --- the
property native Bitcoin already has, extended to a full confidential
DeFi stack.

\section{Fee-aware cryptography}

Witness bytes are the dominant cost. The protocol economizes by picking
the lightest primitive that proves the necessary property:
\begin{itemize}\setlength\itemsep{2pt}
\item \textbf{Bulletproofs+} replaces Bulletproofs for CXFER and AXFER
--- same Pedersen, same kernel sig, $\sim 14\%$ smaller witness across
every transfer\footnote{Chung, H., et al. \emph{Bulletproofs+.} 2020.}.
\item \textbf{Sigma cross-curve binding} (169 bytes, microseconds)
links a secp256k1 chain commitment to a BabyJubJub in-circuit
commitment, avoiding the $\sim 10^{6}$-constraint cost of doing
secp256k1 EC math inside a BN254 SNARK.
\item $\mathrm{T\_SWAP\_VAR}$ \textbf{carries no Groth16.}
Variable-amount fills against a known curve are checkable from public
data; ceremonies are reserved for the surfaces where set-membership
privacy or multi-party amount confidentiality is load-bearing.
\item \textbf{Multi-hop routing} ($\mathrm{T\_SWAP\_ROUTE}$) and
\textbf{batched preauth-take} atomically settle $N$ legs in one Bitcoin
tx, amortizing the $\sim$10\,KB witness floor across many fills.
\item \textbf{Lazy accrual} of LP and farm rewards uses a fixed-point
per-share accumulator (Q.96), so reward crystallization is $O(1)$ per
state-touching event rather than $O(\text{LPs})$, matching
MasterChef\footnote{SushiSwap. \emph{MasterChef contract.} 2020.}.
\end{itemize}

The composition gives smart-contract-shaped properties --- AMM trading,
collateralized wrapping, batched settlement, programmatic yield farms
--- delivered without a VM, without a sidechain, without leaving
Bitcoin L1.

\section{The protocol as one ecosystem}

The primitives fit together as a single system. Any tacit asset works
in any primitive uniformly, so the surfaces below compose into one
self-consistent stack rather than a menu of disconnected products.

\begin{itemize}\setlength\itemsep{2pt}
\item \textbf{Fair-launched native asset.} TAC, 21M-base-unit cap,
deployed via permissionless \texttt{T\_PETCH} / \texttt{T\_PMINT}.
Wire-format-wise a regular tacit asset (amount-private,
shielded-address-compatible, mixer-composable); structurally canonical
to the AMM, bonding, and governance from launch.
\item \textbf{Asset-creation arm.} \texttt{CETCH} issues
confidential-supply assets (optionally mintable);
\texttt{T\_PETCH} $+$ \texttt{T\_PMINT} covers fair-launch capped
issuance with publicly auditable cumulative supply and a height
window; \texttt{T\_DROP} $+$ \texttt{T\_DCLAIM} distributes existing
supply via snapshot-based merkle claims with optional ETH-signature
gating.
\item \textbf{Confidential AMM and yield farms.} Constant-product
pools between any two tacit assets, two trader paths
(\texttt{T\_SWAP\_VAR} cleartext-amount, \texttt{T\_SWAP\_BATCH}
Groth16-private uniform clearing), \texttt{T\_LP\_BOND} /
\texttt{T\_LP\_HARVEST} farm-reward layer over virtual treasuries.
\item \textbf{Cryptographic BTC slots and collateralized fungible
BTC.} cBTC.zk for cryptographic 1:1 wrapping (fixed-denomination
slots), cBTC.tac for fungible BTC-denominated DeFi (amount-granular,
TAC-collateralized) --- both first-class tacit assets that transfer,
swap, mix, and farm like any other.
\item \textbf{Atomic OTC marketplace.} \texttt{T\_AXFER},
\texttt{T\_AXFER\_VAR}, \texttt{T\_SWAP\_ROUTE}, batched preauth-take:
single-tx confidential-asset-vs-BTC settlement with maker/taker
pre-signing that prevents either side from griefing.
\item \textbf{Bonded soft-confirm layer.} \texttt{T\_INTENT\_ATTEST}
preconfirmations, \texttt{T\_WORKER\_SLASH} equivocation slashing on
cryptographic evidence, tacit-mesh cross-worker evidence production.
\item \textbf{Bounded governance.} \texttt{T\_GOV\_*} opcodes for
TAC-weighted parameter votes within explicit safety bands;
load-bearing mechanics remain immutable without a formal amendment.
\end{itemize}

\paragraph{Smart-contract-shaped without a VM.} Smart contracts give
Ethereum users programmable assets, automated market making,
collateralized issuance, atomic multi-leg settlement, yield
distribution, governance, and snapshot-based airdrops. Tacit delivers
the same outcomes by composing the primitives above. Each opcode is a
deterministic state-transition rule any indexer reaches the same
verdict on; the trade-off is real (no arbitrary user-written logic)
and the gains are commensurate (no VM attack surface, no operator
set, no L1 change to add new primitives).

\begin{table}[h]
\centering
\small
\renewcommand{\arraystretch}{1.20}
\begin{tabular}{@{}p{0.32\linewidth} p{0.62\linewidth}@{}}
\toprule
\textbf{Smart-contract pattern} & \textbf{Tacit primitive(s)} \\
\midrule
ERC-20 transfer $+$ optional mint/burn &
\texttt{CETCH} $+$ \texttt{CXFER} $+$ \texttt{T\_MINT} $+$
\texttt{T\_BURN} (Pedersen-committed amounts; mintable if elected at
etch; burn amount is public for auditability). \\
Fair-launch capped permissionless mint &
\texttt{T\_PETCH} $+$ \texttt{T\_PMINT} $+$ \texttt{T\_BURN}
(publicly auditable cap, height window, depth-3 reorg gate). \\
Uniswap V2 AMM &
\texttt{T\_LP\_ADD} / \texttt{T\_LP\_REMOVE} /
\texttt{T\_SWAP\_VAR} / \texttt{T\_SWAP\_BATCH} (virtual pool,
optional per-trader amount confidentiality). \\
MakerDAO CDP &
cBTC.tac via \texttt{T\_CBTC\_TAC\_DEPOSIT} (slot $+$ TAC lien,
$2\times$ cushion, $1.2\times$ force-close, shared insurance pool). \\
Tornado Cash privacy pool &
\texttt{T\_DEPOSIT} / \texttt{T\_WITHDRAW} (Poseidon-Merkle $+$
Groth16 $+$ nullifier set). \\
MasterChef yield farm &
\texttt{T\_FARM\_INIT} / \texttt{T\_LP\_BOND} /
\texttt{T\_LP\_HARVEST} (virtual treasury, Q.96 lazy accrual). \\
Merkle Distributor airdrop &
\texttt{T\_DROP} / \texttt{T\_DCLAIM} (Merkle-snapshot eligibility,
per-claim cap, expiry height). \\
1inch-style multi-leg router &
\texttt{T\_SWAP\_ROUTE} (atomic 2--4-hop AMM routing in one Bitcoin
tx). \\
Optimistic-rollup fraud proof &
\texttt{T\_WORKER\_SLASH} (cryptographic two-attestation evidence,
burn against fixed cap). \\
Compound/Aragon governance &
\texttt{T\_GOV\_PROPOSAL} / \texttt{T\_GOV\_VOTE} /
\texttt{T\_GOV\_VETO} / \texttt{T\_GOV\_EXECUTE} (TAC-weighted,
safety-band-bounded). \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Primitives are immutable; composition is uniform.} Each row
above is an immutable state machine fixed at SPEC time (or
governance-bounded within explicit safety bands). Users compose
primitives; they cannot redefine them. Any tacit asset ---
\texttt{CETCH}'d tokens, LP shares, cBTC.tac, TAC --- transfers via
\texttt{CXFER}, swaps in any pool, deposits into the mixer, and bonds
into farms or worker positions. This is the ERC-20-shape composability
that gave Ethereum DeFi its compounding effect, ported to Bitcoin L1
without a VM.

\paragraph{The stack strengthens with use.} Adoption on any axis lifts
the others: wider TAC use increases shielded-transfer volume and
mixer anonymity sets; deeper AMM liquidity tightens cBTC.tac's
TWAP-based collateral signal and widens the channel back to plain
BTC; a larger bonded base against the fixed cap raises the deterrent
for every future bond; richer farm emissions deepen LP positions
that feed the same liquidity. The same indexer determinism and the
same fixed-cap asset underwrite all of it, and the same primitives
compose across all of it. The ecosystem reinforces itself by being
used.

What this misses: arbitrary user-written contracts. A new primitive
requires a new opcode (and a new circuit if cryptography is
load-bearing), shipped as an additive opcode under SPEC §5.5. The bar is higher than
deploying a smart contract --- but the surface area subject to bugs
is the protocol specification, not every user-written script, and the
soundness argument shrinks accordingly. Tacit is an algebra of
primitives, not an execution environment.

\section{Trust model and limits}

The protocol's load-bearing assumptions deserve naming.

\paragraph{Indexer reference implementation.} The trust target at the
indexer layer is the published specification and the open-source
reference implementations that converge on it, the same shape Bitcoin
Core has at L1. Two indexers running the same code over the same bytes
reach the same verdict; achieving \emph{the same code} is a social
process. Contested edge cases (which have occurred for Runes /
Ordinals indexers in production) resolve via SPEC clarification and
cross-implementation test vectors, not by quorum vote. After the beta
phase stabilizes the wire format under market testing, the TAC DAO
(\S12) takes on bounded stewardship of future protocol versions and
reference-implementation maintenance, under the same safety-band
governance that scopes all other parameter changes. Load-bearing
mechanics --- conservation, slashing semantics, cryptographic
primitives --- remain immutable without a formal SPEC amendment.

\paragraph{TAC bootstrap and market reality.} TAC's initial
distribution went via airdrop to members of the Ethereum-rooted zOrg
DeFi DAO --- a protocol that has operated on Ethereum for over a year
--- with snapshot CSVs preserved at
\href{https://github.com/z0r0z/tacit/tree/main/airdrop}{\texttt{airdrop/}}
in the repo. TAC shares have circulated on the open DeFi market since
the airdrop.
The distribution carried real economic cost on both sides: zOrg
eligibility was earned through ETH spending (share purchase, protocol
fees, or LP farming on Ethereum), and airdrop fulfillment paid Bitcoin
fees to broadcast the on-chain envelopes. On Bitcoin, TAC is etched
via the public-mint \texttt{T\_PETCH} / \texttt{T\_PMINT} mechanism
with a fixed 21M-base-unit cap auditable from chain alone. An active
OTC market for TAC operates on the protocol itself --- settling
asset-vs-BTC trades atomically at \texttt{tacit.finance} through
\texttt{T\_AXFER} / \texttt{T\_AXFER\_VAR} and the atomic-intent
flow --- so market liquidity is already in place, not theoretical.

\paragraph{Persistent utility and reflexive collateral.} The protocol's
security gains weight as TAC market depth grows, and the protocol
structures multiple persistent uses to support that growth: bonding
workers against equivocation, weighting bounded governance,
denominating the canonical $(\mathrm{cBTC.zk}, \mathrm{TAC})$ and
$(\mathrm{cBTC.tac}, \mathrm{TAC})$ AMM pools that route every
BTC-paired trade, receiving protocol-fee revenue from those pools
(\S8), paying out yield-farm rewards, and circulating as a confidential
native asset with full shielded-amount and shielded-address support.
The collateral is partially \textbf{reflexive} --- a stress event that
strains cBTC.tac can also strain TAC --- and the 2.0$\times$ cushion,
1.2$\times$ force-close, and shared insurance pool bound the damage.
The non-bond uses above mitigate the reflexivity by spreading TAC's
value across multiple correlated demand sources; only its open-market
trading and standalone confidential-asset utility are truly exogenous
to cBTC.tac stress.

\paragraph{Network effects on privacy.} TAC's confidential-asset
properties give adoption a self-reinforcing component at every privacy
layer, and the protocol enters this loop with an active wallet base in
the thousands already transacting through TAC's shielded primitives. Every TAC transfer is amount-hidden by default and any receipt
can land at a per-transaction-unique shielded-address marker, so wider
adoption directly increases the volume of shielded activity on chain
--- more cover traffic for an observer to sift through. Mixer use
compounds this further: anonymity sets in TAC mixer pools grow as a
fraction of holders route through the shielded path. Deeper liquidity
in the canonical $(\mathrm{cBTC.zk}, \mathrm{TAC})$ and $(\mathrm{cBTC.tac},
\mathrm{TAC})$ pools widens the channel between confidential balances
and plain BTC sats --- exits to Bitcoin clear larger sums with less
price impact and, when routed through the mixer, with stronger
unlinkability. cBTC.tac's TWAP-based collateral check benefits from the
same depth. Economic-security gains and privacy gains move together at
every layer of the stack.

\paragraph{Bitcoin L1 fee regime.} A confidential transfer carries
${\sim}10\,\text{KB}$ witness, settling at ${\sim}2{,}500$--$3{,}000$
vBytes after the SegWit discount. Tacit is a high-value-transfer
instrument, not a low-value payments rail. Bulletproofs+ ($\sim 14\%$
smaller) and multi-leg routing (\texttt{T\_SWAP\_ROUTE}, batched
preauth-take) amortize the cost across more fills, and Bitcoin
covenant upgrades (\S14) compress further, but no design choice
eliminates the L1 witness floor.

\paragraph{Reorg discipline.} Per-pool deposits, fair-launch mints, and
AMM operations credit only at Bitcoin confirmation depth $\geq 3$ ---
a transient L1 spend that gets reorged out is rolled back along with
its tacit-asset effect. cBTC.zk slot spends are confirmed by Bitcoin
itself; the indexer-tracked nullifier set rolls back with the chain, so
the same $r_{\text{leaf}}$ is re-usable after a reorg iff the
corresponding L1 spend also reorged out. Worker slashes follow the same
rule: an evidence transaction reorged out reverses the slash, and the
dapp's mesh continues to surface the original conflict regardless.

\paragraph{Cross-batch curation MEV.} \texttt{T\_SWAP\_BATCH}'s uniform
clearing price defeats \emph{intra-batch} sandwich attacks by
construction. Cross-batch exclusion --- a settler omitting intents from
a batch they assemble --- is currently bounded by tip economics,
\texttt{T\_INTENT\_ATTEST} preconf-equivocation evidence, and arbitrage
realignment in the next batch. The reserved
\texttt{T\_EXCLUSION\_CLAIM} opcode (drafted) closes this surface in a
follow-up amendment via consensus-level slashing of demonstrable
exclusion.

\paragraph{Ceremony surface.} The mixer's \texttt{withdraw.circom}
runs on one global Phase 2 ceremony (canonical bundle content-addressed
and hardcoded in the dapp); cBTC.zk reuses it without modification. The
AMM's three sub-circuits share Phase 1 (\texttt{pot18}) and run
independent Phase 2 chains anchored to one Bitcoin-block beacon at
finalization --- a single ceremony for the AMM, not a per-pool one.
The proof system's zero-knowledge property protects witness privacy
unconditionally; soundness depends on $\geq 1$ honest contributor
across the ceremony's participants.

\paragraph{Composable compliance.} The mixer pool composes with
association-set techniques (the Privacy Pools construction for opt-in
transparency to known-good UTXO sets) without modifying the protocol
--- an association-set indexer is an additive client-side filter on
the same on-chain envelope set. Tacit ships the cryptographic
primitive; downstream operators choose the policy.

\paragraph{Threat-model summary.} In scope: indexer-spec correctness
(auditable and reproducible from chain), Bitcoin consensus, Groth16 /
Pedersen / Schnorr soundness under standard assumptions, and rational
worker and liquidator behavior. Out of scope: state-level adversaries
that compromise Bitcoin itself, dominant reference-implementation
collusion that violates the SPEC, and tail-correlated TAC/BTC collapse
beyond the bounded-defense surface above. The cryptographic claims hold
under standard assumptions; the economic claims hold conditional on TAC
market depth and an active liquidator population.

\section{Upgrade path}

Tacit is built to absorb improvements in three independent dimensions:
new Bitcoin opcodes, new cryptography, and adjacent privacy protocols.
Each new tacit opcode is purely additive: indexers that don't yet
recognize it skip the envelope; indexers that do recognize it apply
the new rule (SPEC §5.5 unknown-opcode rule). Upgrades land without
breaking existing state or requiring users to migrate value.

\paragraph{Bitcoin covenants.} A covenant-style primitive --- a
CTV-style template check, an \texttt{OP\_CAT}-enabled introspection
construction, an \texttt{OP\_VAULT}-successor design, or a TLUV-style
equivalent --- activates several already-reserved opcodes and tightens
existing ones:
\begin{itemize}\setlength\itemsep{2pt}
\item \textbf{Trustless fractional cBTC.zk}
(\texttt{T\_SLOT\_FRACTIONALIZE} / \texttt{T\_SLOT\_RECONSOLIDATE},
opcodes $\mathtt{0x4D}$ / $\mathtt{0x4E}$, drafted) splits a cBTC.zk
slot into fungible shares and recombines them without a TAC bond. The
cryptographic spec is complete; only the Bitcoin-side enforcement is
missing. cBTC.tac remains for users who prefer the economic-collateral
path; the fractional opcodes become the structurally-trustless
fungible BTC path.
\item \textbf{Aggregated cBTC.zk mixing.} Multiple slots share one
Bitcoin UTXO, giving cBTC.zk slot operations real
BTC-chain-graph privacy on top of the protocol-tree privacy they
already have (\S4).
\item \textbf{On-chain bid escrow.} Today's bid layer is off-chain
because Bitcoin script cannot bind a buyer's pre-signed sats input to
an unknown future asset outpoint; a CTV-style template check closes
that gap, removing one of the last off-chain trust surfaces.
\item \textbf{Covenant-restricted swap inputs} close the residual
intent-race window that currently relies on $\sim$3-block depth
gating in the AMM.
\end{itemize}

\paragraph{ZK in Bitcoin script.} Recent work on Shielded
CSV\footnote{Nick et al., 2024, ``Shielded CSV: Private and Efficient
Client-Side Validation,'' and the broader OP\_CAT-enabled SNARK
verifier line of research.} and OP\_CAT-enabled SNARK verifiers would
let some validation move from the indexer to Bitcoin consensus
itself. Tacit's Groth16 stack is already aligned:
\texttt{withdraw.circom}'s verifying key is content-addressed and
ceremony-anchored, so a proof a worker accepts today verifies under a
script-level checker tomorrow without re-running the ceremony.

\paragraph{Schnorr improvements.} MuSig2 / FROST give multisig
mint-authority with no wire-format change --- the
\texttt{mint\_authority} field is just a 32-byte x-only pubkey.
Cross-Input Signature Aggregation (CISA), if it lands at consensus,
would shrink batched settlements where each fill carries its own
signature today.

\paragraph{Interim privacy composition.} Before covenants land, two
neighboring techniques compose with tacit's existing surfaces:

\begin{itemize}\setlength\itemsep{2pt}
\item \textbf{CoinJoin upstream of mixer deposits or cBTC.zk mints.}
A Bitcoin-layer CoinJoin breaks the fee-source-UTXO clustering that
operationally weakens self-mix unlinkability --- a wallet that funds
its deposit from a CoinJoin output has no chain-graph trail back to
its previous identity.
\item \textbf{BIP-352 silent payments for the plain-BTC funding leg.}
Silent payments compose orthogonally with the shielded address: tacit
hides the recipient marker on tacit-asset receipts; BIP-352 hides it
on plain-sats receipts. Tacit defers a tacit-flavored stealth-sats
scheme rather than competing with the BIP-352 standard, and uses the
mixer pool for the same use case until BIP-352 wallet adoption
converges.
\end{itemize}

\paragraph{Why additivity holds.} Cryptographic primitives replace in
place --- Bulletproofs $\to$ Bulletproofs+ shipped this way, same
Pedersen, same kernel sig, $\sim 14\%$ smaller witness. New Bitcoin
primitives add cryptographic \emph{options} alongside existing economic
ones rather than displacing them: covenants enable fractional cBTC.zk
alongside cBTC.tac, so users can choose either trust model. TAC's
structural roles persist and compound as the protocol surface grows ---
bond layer for workers, weight in governance, denominator and unit of
account in canonical AMM pools (with optional protocol-fee accrual),
reward asset for yield farms, and a confidential native asset in its
own right (shielded amounts and addresses, mixer-composable,
privkey-recoverable). New surfaces --- additional pool families,
stablecoin variants, channel-layer service economies --- broaden the
role set without changing what already ships. The indexer-validated
meta-protocol pattern is a substrate that carries more powerful
primitives as Bitcoin gains them, without ever asking users to migrate
value out and back in.

\section{Conclusion}

We have proposed a self-custodial DeFi layer on Bitcoin L1, built from
a composition of primitives. Two Groth16 circuit families --- anonymous
unique-spend (\texttt{withdraw.circom}) and amount confidentiality
(AMM) --- compose across every privacy-bearing surface. The indexer-validated meta-protocol
pattern that already underwrites Runes and Ordinals is extended into a
\textbf{collateral substrate}: TAC, a fixed 21M-base-unit
asset, bonds workers against equivocation and backs fungible cBTC.tac;
cBTC.zk's ``one leaf, two locks'' construction gives trustless
fixed-denomination BTC slots without a federation. Shielded amounts hide on-chain
values, shielded addresses break recipient clustering at the output
script, and the mixer breaks deposit--withdrawal linkage at the
protocol-tree leaf --- composed at the user's choice of granularity,
the three deliver full privacy across the tacit-asset surface. Merchants opt in only to
shielded amounts and keep a public identity for accounting; payroll
adds shielded addresses for unlinkable per-employee receipts;
privacy-conscious users route through the mixer for unlinkability at
every endpoint within the chosen anonymity set. Wallets reconstruct from the private key and Bitcoin
chain alone.

The cryptographic claims hold under standard assumptions. The economic
claims hold conditional on TAC market depth and an active liquidator
population; limits are stated explicitly in \S13. Every load-bearing
property is either cryptographically enforced from chain alone, or
bounded by an economic mechanism the protocol states in the open.
Where Bitcoin already provides a primitive, tacit uses it. Where Bitcoin
does not, tacit composes indexer-validated state and a fixed-cap bonded
asset to fill the gap. As Bitcoin acquires more script primitives, the
cryptographic surface widens; TAC continues to bond workers, weight
governance, denominate canonical AMM pools (with protocol-fee accrual),
reward yield farms, and carry its own confidential value --- the
structural roles compound as the protocol surface grows (\S14).

\vspace{1em}
\noindent\rule{\linewidth}{0.4pt}
\vspace{0.5em}

\noindent\footnotesize
\emph{Canonical specs:} \texttt{SPEC.md}, \texttt{AMM.md},
\texttt{MIXER.md}, \texttt{spec/CIRCUITS.md},
\texttt{spec/amendments/}.

\end{document}
